---
title: "hw02-Josia-Yuan"
author: "Josia Yuan"
date: "September 28, 2017"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(printr)
library(readr)
library(dplyr)
```

##2)Importing Data

```{r}
#importing using read.csv
cls = c("character","character","factor","character","double")
t = rep("integer",19)
cls = c(cls,t)
dat <- read.csv('../hw02/data/nba2017-player-statistics.csv', stringsAsFactors = FALSE, colClasses = cls)
str(dat)
head(dat)
typeof(dat$Salary)
class(dat$Position)
class(dat$Experience)
```

```{r}
#imporiting using read_cvs9
clt = list(col_character(),col_character(),col_character(),col_character(),col_double(), col_integer(), col_integer(),col_integer(),col_integer(), col_integer(), col_integer(),col_integer(),col_integer(), col_integer(), col_integer(),col_integer(),col_integer(), col_integer(), col_integer(),col_integer(),col_integer(), col_integer(), col_integer(),col_integer())
dat1 <- read_csv('../hw02/data/nba2017-player-statistics.csv', col_types = clt)
dat1$Position <- as.factor(dat1$Position)
str(dat1)
class(dat1$Position)
typeof(dat$Salary)
```

##3)Right after importing data

```{r}
#convert R to 0 in Experience
typeof(dat$Experience)
dat$Experience[dat$Experience == "R"] <- 0
dat$Experience <- as.integer(dat$Experience)
typeof(dat$Experience)
```


##4)Performance of Players

```{r}
#Create variables needed for EFF
dat$Missed_FG <- dat$FGA-dat$FGM
dat$Missed_FT <- dat$FTA-dat$FTM
dat$PTS <- (dat$Points2 * 2)+(dat$Points3 * 3)+dat$FTM
dat$REB <- dat$OREB + dat$DREB
dat$MPG <- dat$MIN/dat$GP
head(dat,5)
```


```{r}
#Calculate EFF
dat$EFF <- (dat$PTS + dat$REB + dat$AST + dat$STL + dat$BLK - dat$Missed_FG - dat$Missed_FT - dat$TO)/dat$GP
head(dat,5)
summary(dat$EFF)
hist(dat$EFF, main='Histogram of Efficiency(EFF)', xlab='EFF')
```



```{r}
#display info of top-10 players by EFF in decreasing order in data frame
top_10 <- arrange(dat,desc(dat$EFF))
head(top_10,10)

#Names of Players with negative EFF
dat[dat$EFF<0,1]

#compute correlation coefficients between EFF and other variables
a<-cor(dat$EFF, dat$PTS)
b<-cor(dat$EFF, dat$REB)
c<-cor(dat$EFF, dat$STL)
d<-cor(dat$EFF, dat$AST)
e<-cor(dat$EFF,dat$BLK)
f<-cor(dat$EFF,dat$Missed_FT)*(-1)
g<-cor(dat$EFF,dat$Missed_FG)*(-1)
h<-cor(dat$EFF,dat$TO)*(-1)
i<-cor(dat$EFF,dat$GP)
#Display cc in descending order
Cor_Coe <- c(a,b,c,d,e,f,g,h,i)
variables <- c('PTS','REB','STL','AST','BLK','MFT','MFG','TO','GP')
R <- data.frame(Variable_with_EFF = variables,
           Correlation_Coefficient = Cor_Coe)
arrange(R,desc(Correlation_Coefficient))
#barchart
Cor_Coe1 <- c(a,b,c,d,e,f,g,h)
variables1<-c('PTS','REB','STL','AST','BLK','MFT','MFG','TO')
color <- c(1,1,1,1,1,2,2,2)
barplot(Cor_Coe1, main = 'Correlations between Player Stats and EFF', names.arg = variables1, ylim = c(-1,1),col=color)
abline(h=0)
```

##5)Efficiency and Salary

```{r}
#scatterplot
plot(dat$EFF,dat$Salary, main='Scatterplot of Efficiency and Salary',xlab='EFF',ylab='Salary')
lines(lowess(dat$EFF,dat$Salary))
text(28,2.5*10^7,'lowess')
#correlation coefficient
cor(dat$EFF,dat$Salary)
```

###Comment:
With a correlation coefficient **0.655624**,the Efficiency of a player doesn't seem to have that strong a correlation with salary.Yet it can still be seen from the scatterplot that a positive correlation is noticable.

```{r}
#select players with MPG>=20
players2<-data.frame(dat[dat$MPG>=20,])
head(players2)
#scatterplot between EFF and Salary
plot(players2$EFF,players2$Salary,main='Scatterplot of EEF and Salary among Players with MPG>= 20',xlab='EFF',ylab='Salary')
lines(lowess(players2$EFF,players2$Salary))
text(29,2*10^7,'lowess')
#correlation coefficient
cor(players2$EFF,players2$Salary)
```

###Comments
With a correlation coefficient **0.5367224**,the Efficiency of "more established" player has an weaker correaltion with salary than that in general. As it can be seen from the scatterplot, points show a tendency of floting around, and the lowess line is less of a stright or soooth one.

##6)Comments and Reflection
Since I'm more familiar with differet functions and arguments used in manipulating data frames, the assignment this time is not as confusing as the first homework. However, I still find it worrisome that I always have to remember to use different sytax for different functions, even though the functions may have the same outcomes.
